A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-08 paper.Rmd
section 2.2 revised

Table \ref{tab:example-residual-plot-table} shows the RESET, Breusch-Pagan and Shapiro-Wilk test results ($p$-values) for the residual plots shown in Figure \ref{fig:residual-plot-common-departures}. The RESET test  and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from R package `lmtest`, respectively. The Shapiro-Wilk test  was computed using the `shapiro.test` from the core R package `stats`. (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) The RESET test is exact, but the power parameter must be chosen. @ramsey_tests_1969 suggests using a power of four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999]. With a sufficiently large sample size, residual-based tests may reject when there is only a slight departure. This is a situation that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance. <!--These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures. -->
 XXX check heteroscedastic spelling

>>>>

Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}. The RESET test and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The Shapiro-Wilk test was computed using the `shapiro.test` from the core R package `stats`. (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter. According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.  The Breusch-Pagan and Shapiro-Wilk tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].

With large sample sizes, hypothesis tests are may reject a residual plot when there is only a slight departure (XXX is there a reference for this statement?). While such rejections may be statistically correct, their sensitivity may render the results impractical. Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses. However, minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes. In fact, the experiment discussed in this paper specifically addresses this.

<!--that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures. -->

### 2023-04-08 paper.tex
section 2.2 revised

Table \ref{tab:example-residual-plot-table} shows the RESET,
Breusch-Pagan and Shapiro-Wilk test results (\(p\)-values) for the
residual plots shown in Figure
\ref{fig:residual-plot-common-departures}. The RESET test and
Breusch-Pagan test were computed using the \texttt{resettest} and
\texttt{bptest} functions from R package \texttt{lmtest}, respectively.
The Shapiro-Wilk test was computed using the \texttt{shapiro.test} from
the core R package \texttt{stats}. (The the R package \texttt{skedastic}
\citep{skedastic} contains a large collection of tests for
heteroskedasticity.) The RESET test is exact, but the power parameter
must be chosen. \citet{ramsey_tests_1969} suggests using a power of
four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests
are approximate.

We would expect the RESET test for non-linearity to reject residual plot
B, the Breusch-Pagan test for heteroskedasticity to reject the residual
plot C, and Shapiro-Wilk test for non-normality to reject residual plot
D, which they all do and correctly fail to reject residual plot A.
Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the
residual plots exhibiting structure that they weren't designed for.
\citet{cook1982residuals} discusses that most residual-based tests for
particular types of departure from model assumptions are also sensitive
to other types of departures, that is, correctly rejected but for the
wrong reason, a phenomenon known as the ``Type III error''. Also, some
types of departure can also have elements of other types of departure,
for example, non-linearity could be viewed as heteroskedasticity.
Additionally, other problesms such as outliers can trigger the rejection
\citep{cook_applied_1999}. With a sufficiently large sample size,
residual-based tests may reject when there is only a slight departure.
This is a situation that is addressed by the results in this paper
because from a practical perspective we would only want to flag the
model as problematic when ther is a XXX diagnostic practical
significance. XXX check heteroscedastic spelling

>>>>

Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the
residual plots in Figure \ref{fig:residual-plot-common-departures}. The
RESET test and Breusch-Pagan test were computed using the
\texttt{resettest} and \texttt{bptest} functions from the R package
\texttt{lmtest}, respectively. The Shapiro-Wilk test was computed using
the \texttt{shapiro.test} from the core R package \texttt{stats}. (The
the R package \texttt{skedastic} \citep{skedastic} contains a large
collection of tests for heteroskedasticity.) Although, the RESET test is
exact, it requires the selection of a power parameter. According to
\citet{ramsey_tests_1969}, a power of four is recommended, which we
adopted in our analysis. The Breusch-Pagan and Shapiro-Wilk tests are
approximate.

We would expect the RESET test for non-linearity to reject residual plot
B, the Breusch-Pagan test for heteroskedasticity to reject the residual
plot C, and Shapiro-Wilk test for non-normality to reject residual plot
D, which they all do and correctly fail to reject residual plot A.
Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the
residual plots exhibiting structure that they weren't designed for.
\citet{cook1982residuals} discusses that most residual-based tests for
particular types of departure from model assumptions are also sensitive
to other types of departures, that is, correctly rejected but for the
wrong reason, a phenomenon known as the ``Type III error''. Also, some
types of departure can also have elements of other types of departure,
for example, non-linearity could be viewed as heteroskedasticity.
Additionally, other problesms such as outliers can trigger the rejection
\citep{cook_applied_1999}.

With large sample sizes, hypothesis tests are may reject a residual plot
when there is only a slight departure (XXX is there a reference for this
statement?). While such rejections may be statistically correct, their
sensitivity may render the results impractical. Therefore, a key goal of
residual plot diagnostics is to identify potential issues that could
lead to incorrect conclusions or errors in subsequent analyses. However,
minor defects in the model are unlikely to have a significant impact and
may be best disregarded for practical purposes. In fact, the experiment
discussed in this paper specifically addresses this.
